---
title: "Workshop 4.17 Solution"
author: "Veerasak Kritsanapraphan"
date: "February 15, 2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include = FALSE}
# my global options defined for each code chunk.
knitr::opts_chunk$set(fig.width=7, fig.height=5, warning=FALSE, message=FALSE, comment = '')
```

## Load recommenderlab

``` {r}
library(dplyr)
library(ggplot2)
library(recommenderlab)
```

Some of the preloaded datasets that come with `recommenderlab` for learning and exploring. 

``` {r}
help(package = "recommenderlab")
datasets_available <- data(package = "recommenderlab")
datasets_available$results[,4] # titles
```

We'll work with the already available *Jester* dataset.

``` {r}
data("Jester5k") # loads dataset
class(Jester5k)
jester_r <- Jester5k 
remove(Jester5k)
```

## Recommender Algorithms Available

The recommender algorithms are stored in a registry object called `recommenderRegistry`. We can get a look at the different models based on the different matrix types.

``` {r}
names(recommenderRegistry$get_entries())
```

### Split Data
```{r}
train <- jester_r[1:500]
test <- jester_r[501:502]
```

### Create Hybrid Recommender
```{r}
recom <- HybridRecommender(
  Recommender(train, method = "POPULAR"),
  Recommender(train, method = "RANDOM"),
  Recommender(train, method = "UBCF"),
  weights = c(.2, .3, .5)
)
recom
```

### Prediction
```{r}
as(predict(recom, test), "list")
```


