---
title: "Linear Regression Prediction"
author: "Veerasak Kritsanapraphan"
date: "2/1/2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading Required R packages

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(modelr)
library(broom)
library(dplyr)
library(ggplot2)
```

#### Load the data
```{r}
data("Wage", package = "ISLR")
```

### Data Sampling
```{r}
set.seed(123)
index <- sample(2, nrow(Wage), replace=TRUE, prob=c(0.7,0.3) )
traindata <- Wage[index==1,]
testdata <- Wage[index==2,]
sprintf("Number of Record in Training Dataset is %d" , nrow(traindata))
sprintf("Number of Record in Testing Dataset is %d" , nrow(testdata))
```

### Compute linear regression model:

### Build the model
```{r}
model <- lm(wage ~ age, data = traindata)
```

### Model performance
```{r}
linearmodel <- data.frame(
  name = 'Linear Regression model',
  R2 = rsquare(model, data=testdata),
  RMSE = rmse(model, data=testdata),
  MAE = mae(model, data=testdata),
  P.value = glance(model)$p.value
)
linearmodel
```

```{r}
glance(model) %>%
  dplyr::select(r.squared, adj.r.squared, sigma, AIC, BIC, p.value)
```

### Visualize the data:

```{r}

ggplot(testdata, aes(age, wage) ) +
  geom_point() +
  stat_smooth(method = lm, formula = y ~ x)
```

## Polynomial regression

```{r}
lm(wage ~ age + I(age^2), data = traindata)
```

An alternative simple solution is to use this:

```{r}
lm(wage ~ poly(age, 2, raw = TRUE), data = traindata)
```

The following example computes a sixfth-order polynomial fit:

```{r}
lm(wage ~ poly(age, 6, raw = TRUE), data = traindata) %>%
  summary()
```

### Build the fifth polynomial model

```{r}
model <- lm(wage ~ poly(age, 5, raw = TRUE), data = traindata)
```

### Model performance
```{r}
polymodel <- data.frame(
  name = 'Polynomial Regression model',
  R2 = rsquare(model, data=testdata),
  RMSE = rmse(model, data=testdata),
  MAE = mae(model, data=testdata),
  P.value = glance(model)$p.value
)
polymodel
```

```{r}
glance(model) %>%
  dplyr::select(r.squared, adj.r.squared, sigma, AIC, BIC, p.value)
```

Visualize the fith polynomial regression line as follow:

```{r}
ggplot(testdata, aes(age, wage) ) +
  geom_point() +
  stat_smooth(method = lm, formula = y ~ poly(x, 5, raw = TRUE))
```

## Log transformation

### Build the model
```{r}
model <- lm(wage ~ log(age), data = traindata)
```

### Model performance
```{r}
logmodel <- data.frame(
  name = 'Log Transform Regression model',
  R2 = rsquare(model, data=testdata),
  RMSE = rmse(model, data=testdata),
  MAE = mae(model, data=testdata),
  P.value = glance(model)$p.value
)
logmodel
```

```{r}
glance(model) %>%
  dplyr::select(r.squared, adj.r.squared, sigma, AIC, BIC, p.value)
```

Visualize the data:
```{r}
ggplot(Wage, aes(age, wage) ) +
  geom_point() +
  stat_smooth(method = lm, formula = y ~ log(x))
```

## Spline regression


```{r}
knots <- quantile(traindata$age, p = c(0.25, 0.5, 0.75))
```

We’ll create a model using a cubic spline (degree = 3):

```{r message=FALSE, warning=FALSE}
library(splines)
```
#### Build the model
```{r}
knots <- quantile(traindata$age, p = c(0.25, 0.5, 0.75))
model <- lm (wage ~ bs(age, knots = knots), data = traindata)
```

#### Model performance
```{r}
splinemodel <- data.frame(
  name = 'Splines model',
  R2 = rsquare(model, data=testdata),
  RMSE = rmse(model, data=testdata),
  MAE = mae(model, data=testdata),
  P.value = glance(model)$p.value
)
splinemodel
```

```{r}
glance(model) %>%
  dplyr::select(r.squared, adj.r.squared, sigma, AIC, BIC, p.value)
```

Visualize the cubic spline as follow:

```{r}
ggplot(testdata, aes(age, wage) ) +
  geom_point() +
  stat_smooth(method = lm, formula = y ~ splines::bs(x, df = 3))
```

## Generalized additive models

```{r message=FALSE, warning=FALSE}
library(mgcv)
```

#### Build the model
```{r}
model <- gam(wage ~ s(age), data = traindata)
```

#### Model performance
```{r}
r <- capture.output(gam.check(model))
p <- strsplit(r[12], " ")[[1]][11]

GAMmodel <- data.frame(
  name = 'Generalized additive model',
  R2 = rsquare(model, data=testdata),
  RMSE = rmse(model, data=testdata),
  MAE = mae(model, data=testdata),
  P.value = p
)
GAMmodel
```

```{r}
glance(model) 
```

Visualize the data:

```{r}
ggplot(testdata, aes(age, wage) ) +
  geom_point() +
  stat_smooth(method = gam, formula = y ~ s(x))
```

## Comparing the models

```{r}
totalmodel <- do.call('rbind', list(linearmodel, polymodel, splinemodel, GAMmodel))
totalmodel
```